---
title: "Homework 2"
author: "Bowen Yao"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(janitor)
```

Problem 1
```{r}
subway = read_csv(
  "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
  show_col_types = FALSE
) |>
  clean_names() |>
  select(
    line, station_name, station_latitude, station_longitude,
    route1:route11, entry, vending, entrance_type, ada
  ) |>
  mutate(entry = entry == "YES")

dim(subway)
```
The dataset contains information about subway entrances and exits, including station name, line, coordinates, routes served, entry permission, vending, entrance type, and ADA compliance. I standardized the column names, kept the requested variables, and converted entry to a logical variable. The cleaned dataset has 1868 rows and 19 columns. It is not tidy yet because the routes are spread across multiple columns.

```{r}
subway |>
  distinct(line, station_name) |>
  nrow()

subway |>
  filter(ada) |>
  distinct(line, station_name) |>
  nrow()
```
There are 465 distinct stations, including 84 ADA-compliant stations. 
```{r}
subway |>
  filter(vending == "NO") |>
  summarise(proportion_allowing_entry = mean(entry))
```
Among entrances/exits without vending, approximately 37.70% allow entry.
```{r}
subway_routes = subway |>
  pivot_longer(
    cols = route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name",
    values_transform = list(route_name = as.character),
    values_drop_na = TRUE
  )

subway_routes |>
  filter(route_name == "A") |>
  distinct(line, station_name) |>
  nrow()

subway_routes |>
  filter(route_name == "A", ada) |>
  distinct(line, station_name) |>
  nrow()
```

There are 60 distinct stations serving the A train, of which 17 are ADA compliant.



Problem2
```{r}
mr_trash = read_excel(
  "data/202610 Trash Wheel Collection Data.xlsx",
  sheet = "Mr. Trash Wheel",
  range = cell_limits(c(2, 1), c(NA, 14))
) |>
  clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    year = as.integer(year),
    sports_balls = as.integer(round(sports_balls)),
    trash_wheel = "Mr. Trash Wheel"
  )

professor_trash = read_excel(
  "data/202610 Trash Wheel Collection Data.xlsx",
  sheet = "Professor Trash Wheel",
  range = cell_limits(c(2, 1), c(NA, 13)),
  na = c("", "NA", "dive")
) |>
  clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    year = as.integer(year),
    trash_wheel = "Professor Trash Wheel"
  )

gwynnda_trash = read_excel(
  "data/202610 Trash Wheel Collection Data.xlsx",
  sheet = "Gwynnda the Good Wheel of the W",
  range = cell_limits(c(2, 1), c(NA, 12))
) |>
  clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    year = as.integer(year),
    trash_wheel = "Gwynnda"
  )

trash_wheels = bind_rows(
  mr_trash,
  professor_trash,
  gwynnda_trash
)
```

```{r}
professor_weight = trash_wheels |>
  filter(trash_wheel == "Professor Trash Wheel") |>
  summarise(total_weight = sum(weight_tons, na.rm = TRUE)) |>
  pull(total_weight)

gwynnda_cigarettes = trash_wheels |>
  filter(
    trash_wheel == "Gwynnda",
    month == "June",
    year == 2022
  ) |>
  summarise(total_cigarettes = sum(cigarette_butts, na.rm = TRUE)) |>
  pull(total_cigarettes)
```

I imported the three sheets, standardized column names, and removed
rows without dumpster numbers. I rounded the sports ball counts to
integers and made the year variable consistent across datasets.
I treated the nonnumeric plastic bottle entry "dive" as missing.
I added a trash wheel identifier before combining the datasets.

The dataset now contains `r nrow(trash_wheels)` observations
and `r ncol(trash_wheels)` variables. Each row is for a dumpster
collection record. Key variables include trash wheel, dumpster number,
date, weight in tons, volume in cubic yards, and counts of individual
types of trash. Professor Trash Wheel collected
`r round(professor_weight, 2)` tons of trash. Gwynnda collected
`r format(gwynnda_cigarettes, cientific = FALSE)` cigarette butts in June 2022.




