---
title: "Homework 2"
author: "Bowen Yao"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(janitor)
```

Problem 1
```{r}
subway = read_csv(
  "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
  show_col_types = FALSE
) |>
  clean_names() |>
  select(
    line, station_name, station_latitude, station_longitude,
    route1:route11, entry, vending, entrance_type, ada
  ) |>
  mutate(entry = entry == "YES")

dim(subway)
```
The dataset contains information about subway entrances and exits, including station name, line, coordinates, routes served, entry permission, vending, entrance type, and ADA compliance. I standardized the column names, kept the requested variables, and converted entry to a logical variable. The cleaned dataset has 1868 rows and 19 columns. It is not tidy yet because the routes are spread across multiple columns.

```{r}
subway |>
  distinct(line, station_name) |>
  nrow()

subway |>
  filter(ada) |>
  distinct(line, station_name) |>
  nrow()
```
There are 465 distinct stations, including 84 ADA-compliant stations. 
```{r}
subway |>
  filter(vending == "NO") |>
  summarise(proportion_allowing_entry = mean(entry))
```
Among entrances/exits without vending, approximately 37.70% allow entry.
```{r}
subway_routes = subway |>
  pivot_longer(
    cols = route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name",
    values_transform = list(route_name = as.character),
    values_drop_na = TRUE
  )

subway_routes |>
  filter(route_name == "A") |>
  distinct(line, station_name) |>
  nrow()

subway_routes |>
  filter(route_name == "A", ada) |>
  distinct(line, station_name) |>
  nrow()
```

There are 60 distinct stations serving the A train, of which 17 are ADA compliant.



Problem2


